|

### 2021-01-05 03-Chp-Systematic-Review-Tools.Rmd
End of day: -3533

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::html_document2: default
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
documentclass: book
---

@@

```{r setup, include=FALSE}
source("R/doc_options.R")
```

@@

Preprints are copies of academic manuscripts that are posted online in advance of being formally published by an academic journal. 
They represent an important source of scientific literature.
A new software program called `medrxivr` was created as part of this thesis to allow researchers to find preprints related to their research in a transparent and reproducible way. 
Development of this tool was required as part of this thesis, as preprints represent a key source of information needed for the research reported in future chapters.

@@

Development of the package followed the

### Success crtieria

Defining success criteria

@@

### Early version

@@

<!----------------------------------------------------------------------------->
## Package infrastructure

@@

## Installation

@@

A second key use of the medrxivr package:

* `mx_download()` Takes the output from `mx_search()` and retrieves the full text PDF for each record, saving it to a folder specified by the user.

```{r, eval = FALSE}
mx_download(results,        # Object returned by mx_search()
            "pdf/")         # Directory to save PDFs to 
```

<!----------------------------------------------------------------------------->
### Reporting your results

One additional function creates a formatted output table with each search strategy presented on an individual line and the number of records associated with this strategy.

```{r, eval = FALSE}
mx_report(results)
```

This allows users to discover which aspects of their

@@

<!----------------------------------------------------------------------------->
### Shiny app

Part of the key theme of accessibility meant creating a web-application that allowed those without any knowledge of the R programming environment to benefit from it's functionality. 
The app is readily available and has

<!----------------------------------------------------------------------------->
## Reception and future plans

```{r, echo = FALSE}
medstats <- cranlogs::cran_downloads("medrxivr", from = "2019-11-22", to = format(Sys.Date(),"%Y-%m-%d") )

@@

The tool has be well received by the community (as of `r date`, `medrxivr` has been downloaded more than `r medrxivr_downloads` times, and the medRxivr app has been visited more than ????), and several use cases have been reported. It has been used to visualize the growing number of preprints related to the 2019 coronavirus outbreak ^[https://twitter.com/L_Brierley/status/1233109086444695553], perform systematic searches in a number of other systematic reviews,[@noone2020] in addition to forming a platform for an analysis of researcher's differing data-sharing tendencies under two different journal models (preprint server vs formal publication).[CITE]

Following rigorous peer-review, it has been onboarded into the rOpenSci suite of packages, a collection of "carefully vetted, staff- and community-contributed R software tools that lower barriers to working with scientific data sources on the web", and an associated article published in the Journal of Open Source Software.[@mcguinness2020a] The entire review discussion is publicly available and can be viewed online. ^[https://github.com/ropensci/software-review/issues/380]

Lobbying of the cold springs harbour laboratory has been ongoing. This would negate the current need to download a full copy of the relevant preprint database before searching it locally, which is currently the rate limiting step for performing searches (for example, as of December 2020, downloading a copy of the bioRxiv database takes approximately an hour.)

@@

<!----------------------------------------------------------------------------->
## Discussion ^[Some of the arguments presented here are repeated in a paper introducing second R package developed during this thesis, but not directly related to it.[@mcguinness] A summary of this second tool, `robvis` is presented in Appendix \@ref(appendix-robvis).]

@@

While searching of the medRxiv database was crucial for the systematic review element of this thesis presented in Chapter \@ref(sys-rev-heading), there are some important limitations to note here. As key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords).
There is the potential that the cross-section of literature posted on medrxiv would be substantially different from the true grey literature - studies or analysis that are not published for a range of reasons, including results that are not deemed "novel" or are not statistically significant. 
This is because simply lowering the barriers to publication may encourage authors to published "null" results,[CITE] but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[CITE] 
It is likely too early, (and too difficult) to tell whether the popularity of preprint repositories will have any effect of the availability of research that was not considered "publishable" at other venues. 
**Of particular interest to this thesis is whether including unpublished preprints in the systematic review and meta-analysis is worthwhile. To this end, the additional contribution of searching these repositories to the systematic review element of this thesis is explored in Chapter \@ref(sys-rev-heading) [CROSS].**

By implementing the tool described above as both as an R package and a `Shiny` web app, the functionality is available to evidence synthesists with varying levels of ability in R. 
These tools serve as an example of the advantages of “packaging” the R scripts that evidence synthesists often create for personal use.[@wickham2015r] 
In the case of `medrxivr`, it is likely that several other evidence synthesists had written personal scripts scripts that have a similar, or related, functionality - in fact, in the course of its development, one other researcher that has done so was identified (Nicholas Fraser, author of the `rbiorxiv` package).[@rbiorxiv] 
If these scripts continue to be developed in private and never shared or publicised, this leads to inevitable harm to the evidence synthesis community, not only in terms of duplication of time and effort, but also due to lost collaboration opportunities.

@@

[**The next section needs to be reduced and fit a bit better to this Chapter - it was taken from the robvis paper, but the issues it raises are important and potentially relevant here.**]{.correction}

@@

* In this Chapter, I have introduced a new tool, `medrxivr`, for performing complex searches in the medRxiv and bioRxiv preprint repositories.

@@

* I have contrasted `medrxivr` with other other platforms to highlight the added functionality it offers. I have also discussed its reception to date, its place in the broader evidence synthesis in R ecosystem, and a roadmap for its future development.

>>>>

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
  bookdown::html_document2: default
  bookdown::pdf_document2:
    template: templates/brief_template.tex
documentclass: book
---

@@

```{r setup, include=FALSE}
source("R/doc_options.R")
source("R/helper.R")
```

@@

Preprints are copies of academic manuscripts that are posted online in advance of being formally published by an academic journal. 
They represent an important source of scientific literature.
A new software program called `medrxivr` was created to allow researchers to find preprints related to their research in a transparent and reproducible way. 
Development of this tool was an essential part of this thesis, as preprints represent a key source of information needed for the research reported in future chapters.

@@

### Success criteria

@@

### Early versions

@@

<!----------------------------------------------------------------------------->
### Package infrastructure

@@

### Installation

@@

A second key use of `medrxivr` is to download the full text papers for records that are deemed eligible for full-text screening (see Figure \@ref(fig:medrxivr-sr)). `mx_download()` takes the list of records included following title and abstract screening, and saves the PDF for each to a folder specified by the user. This funcationality is similar to the "Find Full Text" feature offered by EndNote.

```{r, eval = FALSE}
mx_download(results,        # Object returned by search, less excluded records
            "pdf/")         # Directory to save PDFs to 
```

<!----------------------------------------------------------------------------->
### Refining your search

An important argument of the `mx_search()` is XXXX, which creates a formatted output table with each search strategy presented on an individual line and the number of records associated with this strategy.

```{r, eval = FALSE}
mx_reporter(results)
```

This allows users to discover which aspects of their search is contributing most to the total results returned.

This is crucial as part of developing a search strategy [@bramer2018], as it allows for the key terms related to each topic to be discovered.

@@

A final part of this tool is the ability to visualise the growth of popularity of a particular topic.

**Need to include figures produced by this function here- use the simple search described above**

<!----------------------------------------------------------------------------->
## Discussion

Intro text

<!----------------------------------------------------------------------------->
### Reception and future plans

```{r, echo = FALSE}
medstats <- cranlogs::cran_downloads("medrxivr", from = "2019-11-22", to = format(Sys.Date(),"%Y-%m-%d"))

@@

The tool has be well received by the community (as of `r date`, `medrxivr` has been downloaded more than `r medrxivr_downloads` times, and the medRxivr app has been visited more than ????), and several use cases have been reported. It has been used to visualize the growing number of preprints related to the 2019 coronavirus outbreak ^[https://twitter.com/L_Brierley/status/1233109086444695553], perform systematic searches in a number of other systematic reviews,[@noone2020] and examine how data-sharing behaviour is affected by different publication policies (preprint server versus journal).[CITE]

Following rigorous peer-review, it has been onboarded into the rOpenSci suite of packages, a collection of "carefully vetted, staff- and community-contributed R software tools that lower barriers to working with scientific data sources on the web", and an associated article published in the Journal of Open Source Software.[@mcguinness2020a] The entire review discussion is publicly available and can be viewed online. ^[https://github.com/ropensci/software-review/issues/380] The tool has also been well recieved by the open-source community, demonstrated by the engagement of other developers in contributing to important functionality and suggesting bug-fixes.

Lobbying of the Cold Springs Harbour Laboratory has been ongoing to develop the API further to allow for direct searching of the database. This would negate the current need to download a full copy of the relevant preprint database before searching it locally, which is currently the rate limiting step for performing searches (for example, as of January 2021, downloading a copy of the bioRxiv database takes approximately an hour).

@@

<!----------------------------------------------------------------------------->
### Role of open source tools in evidence synthesis ^[Some of the arguments presented here are repeated in a paper introducing second R package developed during this thesis, but not directly related to the thesis content.[@mcguinness] A summary of this second tool, `robvis`, which creates publication quality risk-of-bias plots, is presented in Appendix \@ref(appendix-robvis).]

@@

While searching of the medRxiv database was crucial for the systematic review element of this thesis presented in Chapter \@ref(sys-rev-heading), there are some important limitations to note here. As key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords).
There is the potential that the cross-section of literature posted on medrxiv would be substantially different from the true grey literature - studies or analyses that are not published for a range of reasons including results that are not deemed "novel" or are not statistically significant. 
This is because simply lowering the barriers to publication may well encourage authors to published "null" results,[CITE] but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[CITE] 
It is likely too early (and too difficult) to tell whether the popularity of preprint repositories will have any effect of the availability of research that was not considered "publishable" at other venues. 
**Of particular interest to this thesis is whether including unpublished preprints in the systematic review and meta-analysis is worthwhile. To this end, the additional contribution of searching these repositories to the systematic review element of this thesis is explored in Chapter \@ref(sys-rev-heading) [CROSS].**

By implementing the tool described above as both as an R package and a `Shiny` web app, the functionality is available to evidence synthesists with varying levels of ability in R. 
These tools serve as an example of the advantages of “packaging” the R scripts that evidence synthesists often create for personal use.[@wickham2015r] 
In the case of `medrxivr`, it is likely that several other evidence synthesists had written personal scripts scripts that have a similar, or related, functionality - in fact, in the course of developing the tool, I identified one other researcher that has done so (Nicholas Fraser, author of the `rbiorxiv` package).[@rbiorxiv] 
If these useful scripts continue to be developed in private and are never shared or publicised, this will inevitably lead to harm to the evidence synthesis community, not only in terms of duplication of time and effort but also due to lost collaboration opportunities.

@@

```{block type='correction'}
**The next section needs to be reduced and fit a bit better to this Chapter - it was taken from the robvis paper, but the issues it raises are important and potentially relevant here.**
```

@@

* In this Chapter, I have introduced a new tool, `medrxivr`, for performing complex systematic searches of the medRxiv and bioRxiv preprint repositories.

@@

* I have contrasted `medrxivr` with other interfaces to medRxiv/bioRxiv data to highlight the added functionality it offers. I have also discussed its reception to date, its place in the broader evidence synthesis in R ecosystem, and a roadmap for its future development.

### 2021-01-05 05-Chp-CPRD-Analysis.Rmd
End of day: -3533

@@

Potential biases included time varying confounding, selection bias due to censoring on death and 
We performed a Cox  test this si to addd some words.

@@

**Comparing and contrasting between different studies is particularly difficult** because of the impact that the use of differnet code list can have on the analysis [@wilkinson2018a; @mcguinness2019c]

>>>>

### Study design and protocol

We performed a prospective cohort study using data from the CPRD. Our initial sample included all participants registered at a participating practice between 1 January 1995 and 29 February 2016 who had a flag for "research quality" data. All events of interest were identified using predetermined code lists, which are available for inspection (see [Data/code availability](#data-code-avail)).

An *a priori* protocol for this study was published, [@walker2016a] and amendments to this are recorded in Supplementary Materials 1. This study was reported in line with the STROBE Cohort guidelines (Supplementary Table 3).[@vonelm2008]

&nbsp;

### Study Cohort

Participants were included in our study cohort if their record contained any of the following index events: a Read code for a diagnosis of hypercholesterolemia or related condition; a Read code for prescription of a lipid-regulating agent (such as statins); a total cholesterol test result of >4mmol/L; or an LDL-c test result of >2mmol/L.

These index events allowed us to define a population of participants who were either at risk of hypercholesterolemia, as indicated by the elevated total or LDL cholesterol test results, or had already been diagnosed with it, as indicated by a diagnostic code/related prescription. This approach was employed in an attempt to reduce confounding by indication that we would expect to observe in the full cohort, because individuals not prescribed lipid-regulating agents likely be less healthy across a range of variables than those prescribed lipid-regulating agents, leading to a biased association been lipid-regulating agent use and dementia. Conditioning entry into the study into the study on being either “at-risk” or already diagnosed with hypercholesterolemia attempts to mitigate this bias.

The index date for a participant was defined as the date where the first relevant code or test result was recorded on their clinical record, and participants were followed up until the earliest of: an outcome of interest; death; end of follow-up (29 February 2016); or last registration date with their GP practice. Participants were removed from our sample if they were less than 40 years of age, had less than 12 months of "research quality" data, were simultaneously prescribed more than one lipid-regulating agent (due to the difficult of assigning these to a single exposure group), or were diagnosed with dementia before or on the date of the index event.

&nbsp;

### Exposures

We considered seven lipid-regulating drug classes based on groupings in the British National Formulary (BNF)[@wishart2017], namely: statins, fibrates, bile acid sequestrants, ezetimibe, nicotinic acid groups, ezetimibe and statin (representing one treatment containing both drugs, rather than the two classes being prescribed concurrently), and omega-3 fatty acid groups.

A participant’s drug class was assigned based on their first recorded prescription, and any drug switching was ignored in an effort to mimic an intention-to-treat approach. We did however examine how often the initial drug class was stopped (defined as last prescription of the primary class being followed by at least six months of observation), added to (defined as a second drug class being prescribed before the last prescription of the initial class), or switched (defined as a second drug class being prescribed after the last prescription of the initial class).

&nbsp;

### Outcomes

We considered five outcomes as part of this analysis: probable  Alzheimer’s disease, possible Alzheimer’s disease, vascular dementia, other dementia, and a composite all-cause dementia outcome (Supplementary Figure 1). When two or more outcomes were coded in a participant’s clinical record, a decision tree was used to differentiate between them (Supplementary Figure 1). The diagnosis date of the outcome was determined by the first record of a relevant code.

&nbsp;

## Covariates

The analysis was adjusted for a range of baseline covariates including sex, grouped year of entry into the cohort (<2000, 2000-2004, 2005-2009, >2010), Charlson co-morbidity index, Index of Multiple Deprivation (IMD), consultation rate, alcohol (current, former,  never), smoking (current, former,  never), BMI, baseline total cholesterol, and history of cardiovascular disease, coronary bypass surgery, coronary artery disease, peripheral arterial disease, hypertension, chronic kidney disease, and Type 1 and Type 2 diabetes. All
covariates were determined at index and definitions for each can be found in Supplementary Table 1.

&nbsp;

@@

Potential biases included time varying confounding, selection bias due to censoring on death and 
We performed a Cox

@@

**Comparing and contrasting between different studies is particularly difficult** because of the impact that the use of different code list can have on the analysis [@wilkinson2018a; @mcguinness2019c]

In order to

As part of our exploration of the, we chose to explore whether our results were comparable with

As discussed in

## Discussion

## Main findings

Lipid-regulating agents had no effect on probable and possible Alzheimer's when compared with no treatment, but were associated with increased risk of an all-cause dementia, vascular dementia and other dementia diagnosis. The effect observed in each case was driven by the statin subgroup, which included a substantial majority of participants. For the other drug classes, no association was found with any outcome, with two exceptions being that ezetimibe was associated with increased risk of vascular and other dementia, while fibrates were associated with increase risk of all-cause dementia and probable Alzheimer's disease.

&nbsp;

## Comparison to other literature

Much of the existing literature focuses on the association of statins alone with neurodegenerative outcomes, with other lipid-regulating agents being grouped as "non-statin cholesterol-lowering drugs".[@ancelin2012] This echoes the distribution of participants among subgroups in our analysis, with the statin subgroup including almost all participants.

&nbsp;

__Statins and all-cause dementia__

A recent Cochrane Review identified two randomized trials comparing treatment with statins versus non-treatment for the prevention of dementia, only one of which presented information on the incidence of dementia.[@mcguinness2016a] This study (Heart Protection Study) showed no effect of treatment with simvastatin on all-cause dementia risk (OR: 1.00,  95%CI:0.61-1.65),[@heartprotectionstudycollaborativegroup2002] but concerns were raised over the diagnostic criteria used. A meta-analysis of 30 observational studies found a reduced risk of all-cause dementia was associated with statin treatment (RR 0.83, 95%CI: 0.79–0.87).[@poly2020]

These sources of evidence conflict with the findings of our analysis, where statin use was associated with an increased risk of all-cause dementia. However, some of the included studies in the meta-analysis specifically exclude vascular dementia from the definition of all-cause dementia,[@chao2015] which may lead to a artifical protective effect of statins on all-cause dementia

&nbsp;

__Statins and Alzheimer's disease__

Our results are broadly in line with the findings of two distinct approaches examining the effect of statin treatment on subsequent Alzheimer's disease. No randomized trials of statins for the prevention of Alzheimer's disease have been reported, but a recent meta-analysis of 20 observational studies found statins were associated with a reduced risk of Alzheimer's disease (RR 0.69, 95% CI 0.60–0.80), though the reduction was more extreme than observed in our analysis.[@poly2020] In addition, a recent Mendelian randomization study examining the effect of genetic inhibition of HMGCR on Alzheimer’s disease found a small reduction in risk of Alzheimer's disease, comparable in magnitude to our findings, but could not rule out no effect (OR: 0.91, 95%CI: 0.63-1.31).[@williams]

An additional analysis found no difference in effect between lipophilic and hydrophilic statins for the prevention of Alzheimer's disease, consistent with a recent meta-analysis.[@chu2018]

&nbsp;

__Statins and non-Alzheimer's disease dementia__

Much less literature is available on the association between lipid-regulating agents and vascular dementia or other dementia. A recent review found four observational studies examining the association of statins and vascular dementia found no effect (RR:0.93, 95% CI 0.74–1.16).[@poly2020] This contrasts with the increased effect found in our analysis. An additional analysis found that lipophilic statins were more harmful than hydrophilic statins in vascular dementia, potentially due to their ability to cross the blood brain barrier.

&nbsp;

__Other drug classes__

Apart from statins, few studies examining a lipid-regulating agent have been reported. One of the few classes for which data was available were fibrates, which were shown to have no effect on all-cause dementia, [@ancelin2012] inconsistent with our finding of a small increase in all-cause dementia risk in those prescribed a fibrate.

To our knowledge, there is no previous study of the effect of preventative treatment with ezetimibe on any dementia outcome, and so we cannot compare our unexpected finding that treatment with the drug associated with an increased risk of the vascular and other dementia outcomes.

&nbsp;

## Strengths and Limitations

A major strength of our analysis is the size of the included cohort and the length of follow-up that the use of electronic health records allowed. In addition, we followed users and non-users from a common index date, using a time-updating treatment indicator to correctly assign time-at-risk to the exposed and unexposed groups.

However, the findings of our analysis are subject to several limitations. There is a strong possibility of differential misclassification of dementia-related condition based on the exposure, as those with memory complaints are more likely to be classified as vascular dementia than Alzheimer's disease if their medical records contains prescriptions for lipid-regulating agents. Further, there is a potential for non-differential misclassification of the outcome based on the use of electronic health records to identify dementia cases.[@wilkinson2018; @mcguinness2019b]

Our study may be subject to confounding by indication, which occurs when factors that affect whether a participant is exposed also affect their outcome. We attempted to address this by limiting inclusion to those either prescribed or “at risk” of being prescribed, which was determined using an elevated test result. We also adjusted for several additional potential confounding variables. However, the negative control analysis of back pain demonstrated a harmful association with lipid-regulating agent use, indicting that our findings may be biased by residual confounding. Important confounding variables for which we have not adjusted could include genetic factors. A recent preprint of a study in the UK Biobank demonstrated that an Alzheimer’s disease polygenic risk score was associated with an increased risk of unspecified Alzheimer’s and vascular dementia, and also with an increased frequency of self-reported raised cholesterol levels, a diagnosis of hypercholesterolaemia, and a history of taking lipid-regulating agents such as statins or ezetimibe.[@korologou-linden2020] This finding, combined with the potential for differential misclassification between Alzheimer’s disease and vascular dementia, could explain part of the observed association between lipid-regulating agents and vascular dementia.

Finally, there is also the potential for reverse causation in this analysis. Dementia and associated conditions have a long prodromal period, during which preclinial disease could cause indications for the prescription of a lipid-regulating agent.

## Conclusions

We have provided new evidence on the potential repurposing of lipid-regulating agents for the prevention of all-cause dementia, Alzheimer’s disease, vascular dementia, and other dementia. We found use of lipid-regulating agents not associated with probable or possible Alzheimer’s disease, but were associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations were driven by those observed in the statin subgroup, which comprised the majority of participants in our cohort.

We have attempted to account for important sources of bias in our analysis and provide a comparison with other available literature. However, there is a strong potential for unmeasured confounding, misclassification and reverse causation, which raises questions about our findings, in particular the unexpected increase in risk of vascular dementia associated with statin use. Future research should aim to address these potential biases and, while it may be costly in terms of time and resources, a large scale, long-term randomized controlled trial would provide useful additional information on the effect of lipid-regulating agents on the risk of dementia and related outcomes.
